Detecting social information in a dense dataset of infants’ natural visual experience

### 2020-01-29 paper/cogsci/paper.Rmd
small edits

abstract: >
      The faces and hands of infants' caregivers and other social partners offer a rich source of social and causal information that may be critical for infants' cognitive and linguistic development. Previous work using manual annotation strategies and cross-sectional data has found systematic changes in the proportion of faces and hands in the egocentric perspective of young infants. Here, we examine the prevalence of faces and hands in a longitudinal collection of over 1400 headcam videos collected from three children along a span of 6 to 32 months of age—the SAYcam dataset [@SAYcam]. To analyze these naturalistic infant egocentric videos, we first validate the use of a modern convolutional neural network for pose detection (OpenPose) for the detection of faces and hands. We then apply this model to the entire dataset, finding a decrease in the proportion of faces vs. hands across age up to the end of the second year of life [@Fausey2016]. In addition, we find considerable variability in the proportion of faces/hands viewed by children in different locations (e.g., living room vs. kitchen), demonstrating that variability across activity contexts may be a driving factor in the amount of social information that infants’ experience.

@@

# Introduction
Infants are confronted by a blooming, buzzing onslaught of stimuli [@james1890principles] which they must learn to parse to make sense of the world around them. Yet they do not embark on this learning process alone: From as early as 3 months of age, young infants follow overt gaze shifts [@gredeback2008microstructure], and even newborns prefer to look at faces with direct vs. averted gaze [@Farroni2002], despite their limited acuity. As faces are likely to be an important conduit of social information that scaffolds cognitive development, cognitive scientists have long hypothesized that faces are prevalent in the visual experience of young infants.

Yet until recently most hypotheses about infants’ visual experience have gone untested. Though parents and scientists alike have strong intuitions about what infant’s see, even the viewpoint of a walking child is not easily predicted by these intuitions [@clerkin2017; @franchak2011]. By equipping infants and toddlers with head-mounted cameras, researchers have begun to document the infant egocentric perspective on the world. Using these methods, a growing body of work now demonstrates that the viewpoints of very young infants (less than 4 months of age) are indeed dominated by frequent, persistent views of the faces of their caregivers [@Jayaraman2015; @Sugden2014; @Jayaraman2018].

As infants motor and cognitive abilities mature beyond these early months, their perspective on the world becomes even more varied. Crawlers see fewer faces and hands than do walking children [@Sanchez2018; @franchak2017see; @kretch2014] as well as different view of objects [@smith2011not]. Further, as infants learn to use their own hands to act on the world, they seem to focus on manual actions taken by their social partners, as their perspective starts to capture views of hands paired with the objects they are acting on [@Fausey2016]. In turn, caregivers may start to use their hands with more communicative intent, directing infants attention with pointing and gestures to different events or objects during play [@yu2013joint].

Here, we analyze the social information present in the infant visual perspective by analyzing a longitudinal collection of over 1400 headcam videos collected from three children along a span of 6 to 32 months of age—the SAYcam dataset [@SAYcam]. In addition to its size and longitudinal nature, this dataset is more naturalistic than those previously used in two key ways. First, recordings were taken under a large variety of contexts, encompassing infants’ viewpoints during both activities outside and inside the home. Even in other naturalistic datasets, the incredible variety in a typical infant’s experience has been largely underrepresented (see examples in Figure 1; e.g., riding in the car, gardening, watching chickens during a walk, browsing magazines, nursing, brushing teeth).  Second, the head-mounted cameras used in the SAYcam dataset captured a substantially larger field of view than those typically used, allowing a more complete picture of the infant perspective. Head-mounted cameras with a more restricted field of view do still capture where infants are foveating most of the time [@yoshida2008, @smith2015contributions] but may still fail to capture short saccades to either faces or hands in the periphery, as the timescale of head movements is much longer.

With hundreds of hours of footage (>29M frames), this large dataset necessitates a shift to an automated annotation strategy. Indeed, annotation of the frames extracted from egocentric videos has been prohibitively time-consuming, meaning that many frames are typically not inspected, even in the most comprehensive studies. For example, @Fausey2016 collected a total of 143 hours of head-mounted camera footage (15.5 million frames), of which one frame every five seconds was hand annotated (by four coders), totalling 103,383 frames (per coder)—an impressive number of annotations but nonetheless only 0.67% of the collected footage. To address this challenge, we use a modern computer vision model of pose detection to automatically detect the presence of hands and faces from the infant egocentric viewpoint. Specifically, we use OpenPose [@Cao2018openpose], a model optimized for jointly detecting human face, body, hand, and foot keypoints that operates well on scenes including multiple people, even if they are partially-occluded (see Figure 1).

In the following paper, we first describe the dataset and validate the use of this model by comparing face and hand detections to a human-annotated set of 24,000 frames. Next, we report how the proportion of faces and hands changes across age. We then investigate sources of variability in our more naturalistic dataset that may explain differences from prior work, including both the field-of-view of the present cameras as well as a diversity of locations during which videos were recorded.

```{r}
## Read in data 
### Load preprocessed and filtered detections
load('../../data/openpose_detections_filtered_threekids.RData') #d_all
load('../../data/openpose_detections_filtered_cropped_allthreekids.RData') # d_cropped
```

@@

## Dataset
The dataset is described in detail in @SAYcam; we summarize these details here. Children wore Veho Muvi miniature cameras mounted on a custom camping headlamp harness (“headcams”) at least twice weekly, for approximately one hour per recording session. One weekly session was on the same day each week at a roughly constant time of day, while the other(s) were chosen arbitrarily at the participating family’s discretion. At the time of the recording, all three children were in single-child households.  Videos captured by the headcam were 640x480 pixels, and a fisheye lens was attached to the camera to increase the field of view to approximately 109 degrees horizontal x 70 degrees vertical. Videos[^1] with technical errors or that were not taken from the egocentric perspective were excluded from the dataset. While data collection for the third child (Y) is still ongoing, here we analyze `r num_videos` videos, with a total duration of `r num_hours` hours (>29 million frames).
[^1]: All videos are available at https://nyu.databrary.org/volume/564

```{r examples, include = T, fig.env = "figure*", fig.pos = "h", fig.align='center', fig.width=8, fig.height=3.5, set.cap.width=T, num.cols.cap=2, fig.align = "center", fig.cap = "Example frames taken from the dataset, illustrating variability in the infant perspective across different locations. OpenPose detections are shown overlaid on these images."}
examples <- png::readPNG("figs/modifiedOPfigure_v3.png")
grid::grid.raster(examples)

@@

 ## evaluate detectors
 gold_sample <- ketan_gold_out %>%
   mutate(face_eval_ketan = evaluate_detector(face_present_ketan, face_openpose), hand_eval_ketan = evaluate_detector(hand_present_ketan, hand_openpose))

 ## output prfs
 face_performance = return_prf_short(gold_sample$face_eval_ketan)
 hand_performance = return_prf_short(gold_sample$hand_eval_ketan)

```
As has been observed in other studies on automated annotation of headcam data [e.g. @frank2013, @sanchez2018postural, @bambach2015lending], detection tasks that are easy in third-person video can be quite challenging in egocentric videos, due to difficult angles and sizes as well as substantial occlusion. For example, the infant perspective often contains non-canonical viewpoints of faces (e.g., looking up at a caregiver's chin) as well as partially-occluded or oblique viewpoints of both faces and hands.

To evaluate OpenPose's performance, we compared its detections to the manually-annotated gold set of frames, calculating precision (hits / hits + false alarms), recall (hits / hits + misses), and f-score (the harmonic mean of precision and recall).  In our data, for faces, the F-score was `r round(face_performance[3],2)`, with a precision of `r round(face_performance[1],2)` and recall of `r round(face_performance[2],2)`. For hands, the F-score was `r round(hand_performance[3],2)`, with a precision of `r round(hand_performance[1],2)` and recall of `r round(hand_performance[2],2)`. While face and hand detections showed moderately good precision, face detections were overall slightly more accurate than hand detections. Hand detections suffered from fairly low recall, indicating that OpenPose likely underestimated the proportion of hands in the dataset. We suspect that this is in part because children’s own hands were often in view of the camera and unconnected to a pose; this factor is explored in detail in the following analysis section.

We also examined whether the precision, recall, and F-score for hands and faces varied per child as a function of age to see if there was any bias in the OpenPose detections related to age, and did not find substantial variation. Nonetheless, while OpenPose was trained on photographs from the adult perspective, this model still generalized relatively well to the egocentric infant viewpoint with no fine-tuning or post-processing of the detections.

@@

We analyzed the social information in view across the entire dataset, looking specifically at the proportions of faces and hands that were in view for each child. Data from videos were binned according to the age of the child (in weeks) (see Figure 2). First, when examining the younger age ranges (< 24 months) we saw that the proportion of faces in view showed a moderate decrease across this age range, both when analyzing the random 24K frames as well as the entire dataset (see Figure 2A).

While these trends appear somewhat different than those observed in @Fausey2016, note that @Fausey2016 included data from very young infants (from 1-4 months of age), while here the youngest videos coming from S and A around 6 and 9 months of age, respectively.  Similarly, our age range extends 8 months later than those infants in @Fausey2016, throughout a portion of third year of life.

```{r FaceVsHands, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion of faces (A), hands (B), and faces vs. hands (C) in the entire dataset and (D). when hand detections are restricted to the upper 60 percent of the field of view (B). Data from each child are plotted separately and binned by each week that the videos were filmed; data pointsscaled by the number of frames in that age range."}

faces <- ggplot(summary_by_age, aes(x=age_day_bin, y=prop_faces, size=num_detect, col=child_id)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10) + 
  ylab('Proportion Faces') + 
  # theme_few(base_size=16) +
  xlab('Age (months)') +
  ylim(0,.5) +
  theme(legend.position="none") +
  ggtitle('(A).') +
  scale_colour_brewer(palette = "Dark2")

hands <- ggplot(summary_by_age, aes(x=age_day_bin, y=prop_hands, size=num_detect, col=child_id)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10) + 
  # theme_few(base_size=16) +
  ylab('Proportion Hands') + 
  xlab('Age (months)') +
  ylim(0,.5)+
  theme(legend.position="none") +
  ggtitle('(B).') +
  scale_colour_brewer(palette = "Dark2")

face_v_hand <- ggplot(summary_by_age, aes(x=age_day_bin, y=prop_faces - prop_hands, size=num_detect, col=child_id)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10) + 
  # theme_few(base_size=16) +
  ylab('Proportion Faces - Hands') + 
  xlab('Age (months)') +
  ylim(-.3, .3) +
  theme(legend.position = "none") +
  ggtitle('(C).') +
  scale_colour_brewer(palette = "Dark2")

face_v_hand_cropped <- ggplot(summary_by_age_cropped, aes(x=age_day_bin, y=prop_faces - prop_hands, size=num_detect, col=child_id)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10) + 
  # theme_few(base_size=16) +
  ylab('Proportion Faces - Hands') + 
  xlab('Age (months)') +
  ylim(-.3, .3) +
  theme(legend.position = "none") +
  ggtitle('(D).')  +
  scale_colour_brewer(palette = "Dark2")

ggarrange(faces, hands, face_v_hand, face_v_hand_cropped, nrow=1)
```

However, the most striking result is a much greater proportion of hands in view than have previously been reported [@Fausey2016]. We found this to be true across all ages, in both children, and regardless of whether we analyzed human annotations (on the 24K random subset) or OpenPose annotations on entire dataset. This is notable especially given that OpenPose showed relatively low recall for hands, indicating that this is still an underestimate of the proportion of hands in view. One reason this could be the case is the much larger field of view that was captured by the cameras used in this study: unlike previous studies, our cameras were outfitted with a fish-eye lens in an attempt to capture as much of the children’s field of view as possible. Thus, the field of view (FOV) of the fisheye lens used was larger (109 degrees horizontal x 70 degrees vertical) than the FOV of the lens used in many previous studies; for example, in @Fausey2016 the FOV was 69 x 41 degrees. This larger field of view may have allowed the SAYcam cameras to capture not only the presence of a social partner’s hands interacting with objects, but also the children’s own hands, leading to more frequent hand detections.

```{r}
turk_annotations_seg = "../../data/turk_segmentations_hands_only_processed.csv" 
g_seg = read.csv(turk_annotations_seg) %>%
  select(-HITID, -HITTypeId, -WorkerID) %>%
  mutate(center_x = left + width/2, center_y = top + height/2)

adult_hands <- g_seg %>%
  filter(label=="Adult hand")

##
child_hand_plot <- ggplot(child_hands, aes(x=center_x, y=center_y)) +
  geom_point() + 
  stat_density_2d(aes(fill = ..level..), geom="polygon") +
  coord_fixed(ratio=1) +
  # ggtitle('(A). Child hand density') +
  ylim(0,480) +
  xlim(0,640) +
  ylab('') +
  xlab('') +
  theme_few(base_size=12) +
  theme(legend.position="none", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  scale_y_reverse()
##
adult_hand_plot <- ggplot(adult_hands, aes(x=center_x, y=center_y)) +
  geom_point() + 
  stat_density_2d(aes(fill = ..level..), geom="polygon") +
  coord_fixed(ratio=1) +
  ylim(0,480) +
  xlim(0,640) +
  scale_y_reverse() +
  ylab('') +
  xlab('') +
  theme_few(base_size=12) +
    theme(legend.position="none", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) 
  # ggtitle('(B). Adult hand density') 
```

```{r density, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=4, fig.height=2, fig.cap = "Density estimates for the child (left) and adult (right) hands that were detected in teh 24K random gold set; each dot represents the center of a bounding box made by an adult participant."}
ggarrange(child_hand_plot, adult_hand_plot, nrow=1)
```

To assess this possibility, we obtained human annotations for a subsample of 9051 frames in which a hand was detected in the random gold set; participants (recruited via AMT) were asked to draw bounding boxes around children’s hands and adult’s hands. Overall, we found that 34% of the hands in the random 24K sample were of children’s own hands (compared to 8% reported in @Fausey2016) suggesting that this difference in field of view did contribute substantially to the higher rate of hand detections in the frames. Heatmaps of the bounding boxes obtained from these annotations can be seen in Figure 3, showing that children’s hands tended to appear in the lower half of the frames.

We thus re-analyzed the entire dataset while restricting our analysis to a smaller, upper portion of the frame comparable to the field of view used in @Fausey2016. To do so, we excluded hand detections that occurred in the bottom 40% of the frame, while retaining all hand detections that occurred in the top 60% of the frame. This coarse cropping of decreased the proportion of hand detections from `r mean(d_all$hand_openpose)`percent to `r mean(d_cropped$hand_detected_cropped)` percent.

With this modified field of view, we now observed a more substantive decrease across age in the proportions of faces relative to hands. To quantify this trend, we fit a generalized linear mixed model to the binned detection rates from this smaller field of field to estimate changes in the proportion of faces vs. hands over age. To better approximate @Fausey2016, we also restricted our analysis to the age range seen in @Fausey2016, excluding videos when children were over 24 months of age, and indeed found that the proportion of faces vs. hands declined modestly with age (see Table 1).

```{r eval=FALSE}
summary_by_age_days_cropped_limited <- d_cropped %>%
  mutate(age_day_bin = cut(age_days, bins, labels=round(bin_starts/30,1))) %>%
  mutate(age_day_bin = as.numeric(as.character(age_day_bin))) %>%
  group_by(age_day_bin,child_id) %>%
  filter(age_day_bin < (24))  %>%
  summarize(num_faces = sum(face_openpose), num_hands = sum(hand_detected_cropped)) %>%
  rename(age_months = age_day_bin)

summary_by_age_days_cropped <- d_cropped %>%
  mutate(age_day_bin = cut(age_days, bins, labels=round(bin_starts/30,1))) %>%
  mutate(age_day_bin = as.numeric(as.character(age_day_bin))) %>%
  group_by(age_day_bin,child_id) %>%
  summarize(num_faces = sum(face_openpose), num_hands = sum(hand_detected_cropped)) %>%
  rename(age_months = age_day_bin)

>>>>

abstract: >
      The faces and hands of infants' caregivers and other social partners offer a rich source of social and causal information that may be critical for infants' cognitive and linguistic development. Previous work using manual annotation strategies and cross-sectional data has found systematic changes in the proportion of faces and hands in the egocentric perspective of young infants. Here, we examine the prevalence of faces and hands in a longitudinal collection of over 1400 headcam videos collected from three children along a span of 6 to 32 months of age—the SAYcam dataset [@SAYcam]. To analyze these naturalistic infant egocentric videos, we first validate the use of a modern convolutional neural network for pose detection (OpenPose) for the detection of faces and hands. We then apply this model to the entire dataset, finding a decrease in the proportion of faces vs. hands across age up to the end of the second year of life [@Fausey2016]. In addition, we find considerable variability in the proportion of faces/hands viewed by children in different locations (e.g., living room vs. kitchen), demonstrating that variability across activity contexts may be a driving factor in the amount of social information that infants experience.

@@

# Introduction
Infants are confronted by a blooming, buzzing onslaught of stimuli [@james1890principles] which they must learn to parse to make sense of the world around them. Yet they do not embark on this learning process alone: From as early as 3 months of age, young infants follow overt gaze shifts [@gredeback2008microstructure], and even newborns prefer to look at faces with direct vs. averted gaze [@Farroni2002], despite their limited acuity. As faces are likely to be an important conduit of social information that scaffolds cognitive development, psychologists have long hypothesized that faces are prevalent in the visual experience of young infants.

Yet until recently most hypotheses about infants' visual experience have gone untested. Though parents and scientists alike have strong intuitions about what infants see, even the viewpoint of a walking child is not easily predicted by these intuitions [@clerkin2017; @franchak2011]. By equipping infants and toddlers with head-mounted cameras, researchers have begun to document the infant's egocentric perspective on the world. Using these methods, a growing body of work now demonstrates that the viewpoints of very young infants (less than 4 months of age) are indeed dominated by frequent, persistent views of the faces of their caregivers [@Jayaraman2015; @Sugden2014; @Jayaraman2018].

Beyond these early months,infants motor and cognitive abilities mature, leading to a vastly different perspectives on the world. For example, crawlers see fewer faces and hands than do walking children [@Sanchez2018; @franchak2017see; @kretch2014] as well as different views of objects [@smith2011not]. Further, as infants learn to use their own hands to act on the world, they seem to focus on manual actions taken by their social partners, and their perspective starts to capture views of hands manipulating objects [@Fausey2016]. In turn, caregivers may also start to use their hands with more communicative intent, directing infants attention by pointing and gesturing to different events and objects during play [@yu2013joint].

Here, we examine the social information present in the infant visual perspective--the presence of faces and hands--by analyzing a longitudinal collection of over 1400 headcam videos collected from three children along a span of 6 to 32 months of age—the SAYcam dataset [@SAYcam]. In addition to its size and longitudinal nature, this dataset is more naturalistic than those previously used in two key ways. First, recordings were taken under a large variety of activity contexts [@roy2015predicting; @bruner1985role] encompassing infants' viewpoints during both activities outside and inside the home. Even in other naturalistic datasets, the incredible variety in a typical infan's experience has been largely underrepresented (see examples in Figure 1; e.g., riding in the car, gardening, watching chickens during a walk, browsing magazines, nursing, brushing teeth). Second, the head-mounted cameras used in the SAYcam dataset captured a larger field of view than those typically used, allowing a more complete picture of the infant perspective. While head-mounted cameras with a more restricted field of view do represent where infants are foveating most of the time [@yoshida2008, @smith2015contributions], they may fail to capture short saccades to either faces or hands in the periphery, as the timescale of head movements is much longer.

With hundreds of hours of footage (>37.5M frames), however, this large dataset necessitates a shift to an automated annotation strategy. Indeed, annotation of the frames extracted from egocentric videos has been prohibitively time-consuming, meaning that many frames are typically not inspected, even in the most comprehensive studies. For example, @Fausey2016 collected a total of 143 hours of head-mounted camera footage (15.5 million frames), of which one frame every five seconds was hand-annotated (by four coders), totalling 103,383 frames (per coder)—an impressive number of annotations but nonetheless only 0.67% of the collected footage. To address this challenge, we use a modern computer vision model of pose detection to automatically detect the presence of hands and faces from the infant egocentric viewpoint. Specifically, we use OpenPose [@Cao2018openpose], a model optimized for jointly detecting human face, body, hand, and foot keypoints that operates well on scenes including multiple people, even if they are partially-occluded (see Figure 1).

In the following paper, we first describe the dataset and validate the use of this model by comparing face and hand detections to a human-annotated set of 24,000 frames. Next, we report how the proportion of faces and hands changes with age in each of the three children in the dataset. We then investigate sources of variability in our more naturalistic dataset that may explain differences from prior work, including both the field-of-view of the present cameras as well as a diversity of locations during which videos were recorded.

```{r}
## Read in data 
### Load preprocessed and filtered detections
load('../../data/openpose_detections_filtered_threekids.RData') #d_all
load('../../data/openpose_detections_filtered_cropped_allthreekids_jan28.RData') # d_cropped
```

@@

## Dataset
The dataset is described in detail in @SAYcam; we summarize these details here. Children wore Veho Muvi miniature cameras mounted on a custom camping headlamp harness ("headcams") at least twice weekly, for approximately one hour per recording session. One weekly session was on the same day each week at a roughly constant time of day, while the other(s) were chosen arbitrarily at the participating family’s discretion. At the time of the recording, all three children were in single-child households.  Videos captured by the headcam were 640x480 pixels, and a fisheye lens was attached to the camera to increase the field of view to approximately 109 degrees horizontal x 70 degrees vertical. Videos[^1] with technical errors or that were not taken from the egocentric perspective were excluded from the dataset. While data collection for the third child (Y) is still ongoing, here we analyze `r num_videos` videos, with a total duration of `r round(num_hours,2)` hours (>37.5 million frames).
[^1]: All videos are available at https://nyu.databrary.org/volume/564

```{r examples, include = T, fig.env = "figure*", fig.pos = "h", fig.align='center', fig.width=8, fig.height=3.5, set.cap.width=T, num.cols.cap=2, fig.align = "center", fig.cap = "Example frames taken from the dataset, illustrating variability in the infant perspective across different locations. OpenPose detections are shown overlaid on these images (green dots = face, red dots = hands, orange dots = pose)."}
examples <- png::readPNG("figs/modifiedOPfigure_v3.png")
grid::grid.raster(examples)

@@

 ## evaluate detectors
 gold_sample <- ketan_gold_out %>%
  select(-vid_name) %>%
  rename(vid_name = vid_name_short) %>% 
  mutate(vid_name = as.factor(vid_name)) %>%
  mutate(face_eval_ketan = evaluate_detector(face_present_ketan, face_openpose), hand_eval_ketan = evaluate_detector(hand_present_ketan, hand_openpose))

 ## output prfs
 face_performance = return_prf_short(gold_sample$face_eval_ketan)
 hand_performance = return_prf_short(gold_sample$hand_eval_ketan)
```

```{r} 
### Load turk hand annotations wtih all bounding boxes (multiple dets per frame)
turk_annotations_seg = "../../data/turk_segmentations_hands_only_processed.csv" 
g_seg = read.csv(turk_annotations_seg) %>%
  select(-HITID, -HITTypeId, -WorkerID) %>%
  mutate(center_x = left + width/2, center_y = top + height/2)

adult_hands <- g_seg %>%
  filter(label=="Adult hand")
```

```{r eval=FALSE}
hands_missed <- gold_sample %>%
  select(-vid_name) %>%
  rename(vid_name = vid_name_short) %>% 
  mutate(vid_name = as.factor(vid_name)) %>%
  left_join(child_adult_hand_annotations, by=(c("vid_name","frame"))) %>%
  filter(!is.na(full_image_path)) %>%
  filter(hand_eval_ketan == 'FN')

# render examples of missed hands
# dir.create(paste0('det_examples/missed_hand'), recursive=TRUE)
# for (i in seq(1,length(hands_missed$full_image_path),1)){
#   image_read(as.character(hands_missed$full_image_path[i])) %>%
#     image_append(stack = FALSE) %>%
#     image_write(file.path(paste0("det_examples/missed_hand/", hands_missed$vid_name[i],hands_missed$frame[i],'.jpg')))
# }

```{r}
##
child_hand_plot <- ggplot(child_hands, aes(x=center_x, y=center_y)) +
  geom_point() + 
  stat_density_2d(aes(fill = ..level..), geom="polygon") +
  coord_fixed(ratio=1) +
  # ggtitle('(A). Child hand density') +
  ylim(0,480) +
  xlim(0,640) +
  ylab('') +
  xlab('') +
  theme_few(base_size=12) +
  theme(legend.position="none", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  scale_y_reverse()
##
adult_hand_plot <- ggplot(adult_hands, aes(x=center_x, y=center_y)) +
  geom_point() + 
  stat_density_2d(aes(fill = ..level..), geom="polygon") +
  coord_fixed(ratio=1) +
  ylim(0,480) +
  xlim(0,640) +
  scale_y_reverse() +
  ylab('') +
  xlab('') +
  theme_few(base_size=12) +
    theme(legend.position="none", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) 
  # ggtitle('(B). Adult hand density') 
```

```{r density, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=4, fig.height=2, fig.cap = "Density estimates for the child (left) and adult (right) hands that were detected in the 24K frame random gold set; each dot represents the center of a bounding box made by an adult participant."}
ggarrange(child_hand_plot, adult_hand_plot, nrow=1)
```

```{r} 
gold_sample_no_child_hands <- gold_sample %>%
  left_join(child_adult_hand_annotations) %>%
  replace_na(list(child_hand_seg = FALSE)) %>% # replace NAs so they are counted
  filter(child_hand_seg==FALSE) %>%
  mutate(hand_eval_adults = evaluate_detector(hand_present_ketan, hand_openpose))

hand_performance_adults = return_prf_short(gold_sample_no_child_hands$hand_eval_adults)
```

As has been observed in other studies on automated annotation of headcam data [e.g. @frank2013, @sanchez2018postural, @bambach2015lending], detection tasks that are easy in third-person video can be quite challenging in egocentric videos, due to difficult angles and sizes as well as substantial occlusion. For example, the infant perspective often contains non-canonical viewpoints of faces (e.g., looking up at a caregiver's chin) as well as partially-occluded or oblique viewpoints of both faces and hands. Furthermore, hand detection tends to be a somewhat harder problem than face detection [@simon2017hand; @bambach2015lending], and certainly one that has received less attention than face detection. We thus expected overall performance to be lower in these naturalistic videos than on either photos taken from the adult perspective or in egocentric videos in controlled, laboratory settings [@.g., @sanchez2018postural].

To evaluate OpenPose's performance, we compared its detections to the manually-annotated gold set of frames, calculating precision (hits / hits + false alarms), recall (hits / hits + misses), and f-score (the harmonic mean of precision and recall).  In our data, for faces, the F-score was `r round(face_performance[3],2)`, with a precision of `r round(face_performance[1],2)` and recall of `r round(face_performance[2],2)`. For hands, the F-score was `r round(hand_performance[3],2)`, with a precision of `r round(hand_performance[1],2)` and recall of `r round(hand_performance[2],2)`. While face and hand detections showed moderately good precision, face detections were overall slightly more accurate than hand detections.  In general, hand detections suffered from fairly low recall, indicating that OpenPose likely underestimated the proportion of hands in the dataset.

We suspected that this is in part because children’s own hands were often in view of the camera and unconnected to a pose—a notoriously challenging detection problem [@bambach2015lending]. To assess this possibility, we obtained human annotations for a subsample of 9051 frames in which a hand was detected; participants (recruited via AMT) were asked to draw bounding boxes around children’s hands and adult’s hands. Overall, we found that almost half `r mean(hands_missed$child_hand_seg)` of missed hand detections were of child hands. When frames with children's hands were removed from the gold set, recall did improve to `r hand_performance_adults[3]` somewhat. We also observed that children’s hands tended to appear in the lower half of the frames; heatmaps of the bounding boxes obtained from these annotations can be seen in Figure 3.

Finally, we examined whether the precision, recall, and F-score for hands and faces varied with age or child, and did not find substantial variation. Thus, while OpenPose was trained on photographs from the adult perspective, this model still generalized relatively well to the egocentric infant viewpoint with no fine-tuning or post-processing of the detections.

@@

summary_by_age_cropped <- d_cropped %>%
  replace_na(list(hand_detected = FALSE, face_detected = FALSE)) %>%
  mutate(age_day_bin = cut(age_days, bins, labels=round(bin_starts/30,1))) %>%
  mutate(age_day_bin = as.numeric(as.character(age_day_bin))) %>%
  group_by(age_day_bin,child_id) %>%
  summarize(num_detect = length(face_openpose), prop_faces = sum(face_detected) / num_detect,  prop_hands = sum(hand_detected) / num_detect) 
```

```{r FaceVsHands, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=4, fig.cap = "Proportion of faces (A), hands (B), and faces vs. hands (C) in the entire dataset and (D). when hand detections are restricted to the upper 60 percent of the field of view (B). Data from each child are plotted separately and binned by each week that the videos were filmed; data pointsscaled by the number of frames in that age range."}

faces <- ggplot(summary_by_age, aes(x=age_day_bin, y=prop_faces, size=num_detect, col=child_id)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10) + 
  ylab('Prop. Faces') + 
  # theme_few(base_size=16) +
  xlab('Age (months)') +
  ylim(0,.5) +
  theme(legend.position="none") +
  ggtitle('A.') +
  scale_colour_brewer(palette = "Dark2")

hands <- ggplot(summary_by_age, aes(x=age_day_bin, y=prop_hands, size=num_detect, col=child_id)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10) + 
  # theme_few(base_size=16) +
  ylab('Prop. Hands') + 
  xlab('Age (months)') +
  ylim(0,.5)+
  theme(legend.position="none") +
  ggtitle('B.') +
  scale_colour_brewer(palette = "Dark2")

face_v_hand <- ggplot(summary_by_age, aes(x=age_day_bin, y=prop_faces - prop_hands, size=num_detect, col=child_id)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10) + 
  # theme_few(base_size=16) +
  ylab('Prop. Faces - Hands') + 
  xlab('Age (months)') +
  ylim(-.3, .3) +
  theme(legend.position = "none") +
  ggtitle('C.') +
  scale_colour_brewer(palette = "Dark2")
###

faces_cropped <- ggplot(summary_by_age_cropped, aes(x=age_day_bin, y=prop_faces, size=num_detect, col=child_id)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10) + 
  ylab('Prop. Faces (Upper FOV)') + 
  # theme_few(base_size=16) +
  xlab('Age (months)') +
  ylim(0,.5) +
  theme(legend.position="none") +
  ggtitle('D.') +
  scale_colour_brewer(palette = "Dark2")

hands_cropped <- ggplot(summary_by_age_cropped, aes(x=age_day_bin, y=prop_hands, size=num_detect, col=child_id)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10) + 
  # theme_few(base_size=16) +
  ylab('Prop. Hands (Upper FOV)') + 
  xlab('Age (months)') +
  ylim(0,.5)+
  theme(legend.position="none") +
  ggtitle('E.') +
  scale_colour_brewer(palette = "Dark2")

face_v_hand_cropped <- ggplot(summary_by_age_cropped, aes(x=age_day_bin, y=prop_faces - prop_hands, size=num_detect, col=child_id)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10) + 
  # theme_few(base_size=16) +
  ylab('Prop. Faces - Hands (Upper FOV)') + 
  xlab('Age (months)') +
  ylim(-.3, .3) +
  theme(legend.position = "none") +
  ggtitle('F.')  +
  scale_colour_brewer(palette = "Dark2")

ggarrange(faces, hands, face_v_hand, faces_cropped, hands_cropped, face_v_hand_cropped, nrow=2)
```

We analyzed the social information in view across the entire dataset, looking specifically at the proportions of faces and hands that were in view for each child. Data from videos were binned according to the age of the child (in weeks) (see Figure 2). First, when examining the younger age ranges (< 24 months), as in @Fausey2016, we saw that the proportion of faces vs. hands in view showed a moderate decrease across this age range, both when analyzing the random 24K frames as well as the entire dataset (see Figure 2A).

However, the most striking result is a much greater proportion of hands in view than have previously been reported [@Fausey2016]. We found this to be true across all ages, in all three children, and regardless of whether we analyzed human annotations (on the 24K random subset) or OpenPose annotations on the entire dataset. This is notable especially given that OpenPose showed relatively low recall for hands, indicating that this is still an underestimate of the proportion of hands in view. Nonetheless, one reason this could be the case is the much larger field of view that was captured by the cameras used in this study: unlike previous studies, our cameras were outfitted with a fish-eye lens in an attempt to capture as much of the children’s field of view as possible. Thus, the field of view (FOV) of the fisheye lens used was larger (109 degrees horizontal x 70 degrees vertical) than the FOV of the lens used in many previous studies; for example, in @Fausey2016 the FOV was 69 x 41 degrees. This larger field of view may have allowed the SAYcam cameras to capture not only the presence of a social partner’s hands interacting with objects, but also the children’s own hands, leading to these more frequent hand detections.

We thus re-analyzed the entire dataset while restricting our analysis to a smaller, upper portion of the frame comparable to the field of view used in @Fausey2016. To do so, we excluded hand detections that occurred in the bottom 40% of the frame, while retaining all hand detections that occurred in the top 60% of the frame. This coarse cropping of decreased the proportion of hand detections from `r round(mean(d_all$hand_openpose),2)`percent to `r round(mean(d_cropped$hand_detected_cropped),2)` percent, but only decreased face detections from `r round(mean(d_all$face_openpose),2)` to `r round(mean(d_cropped$face_detected_cropped,2)`.  Within this modified field of view, we still observed, on average, more hand detections than face detections.

However, we also still observed a moderate decrease in the proportion of faces vs. hands in view across age. To quantify this trend, we fit a generalized linear mixed model to the binned detection rates from this smaller field of field to estimate changes in the proportion of faces vs. hands over age. To better approximate @Fausey2016, we  restricted our analysis to the age range seen in @Fausey2016, excluding videos when children were over 24 months of age, and indeed found that the proportion of faces vs. hands declined modestly with age (see Table 1). Note that @Fausey2016 also included data from very young infants (from 1-4 months of age), while here the youngest videos coming from S and A around 6 and 9 months of age, respectively.

```{r eval=FALSE}
summary_by_age_days_cropped_younger <- d_cropped %>%
  replace_na(list(hand_detected = FALSE, face_detected = FALSE)) %>%
  filter(age_days < 730) %>% ## less than 24 months
  group_by(age_days,child_id) %>%
  summarize(num_faces = sum(face_openpose), num_hands = sum(hand_detected), num_detect = length(face_openpose)) %>%
  mutate(prop_faces = num_faces / num_detect, prop_hands = num_hands / num_detect)

summary_by_age_days_cropped <- d_cropped %>%
  replace_na(list(hand_detected = FALSE, face_detected = FALSE)) %>%
  group_by(age_days,child_id) %>%
  summarize(num_faces = sum(face_openpose), num_hands = sum(hand_detected), num_detect = length(face_openpose)) %>%
  mutate(prop_faces = num_faces / num_detect, prop_hands = num_hands / num_detect)
